TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages

### 2021-03-31 article.tex
TTT: footnotes en español

\en{The normalized posterior is found by dividing the right hand by the evidence $P(r)$.}
\es{Donde el el posterior normalizado se obtiene dividiendo el lado derecho por la evidencia, $P(r)$.}

@@

\es{Es interesante notar las similitudes y diferencias entre el likelihood y la evidencia.}

@@

\en{In this last example we use the entire history of matches of the Association of Tennis Professionals (ATP).}
\es{En este último ejemplo utilizamos toda la historia de partidas de la Asociación de Tenistas Profesionales (ATP).}
%
\en{The database has \num{447000} games starting at year 1915 until 2020 with more than \num{19000} participating players and is publicly available\footnote{Source data at \nolinkurl{https://github.com/JeffSackmann/tennis_atp}}.}
\es{La base de datos cuenta con \num{447000} mil partidas que van desde el año 1915 hasta el 2020, en las que participan \num{19000} jugadores.}
%
\en{The information is formatted and stored in a single \texttt{csv} file\footnote{Available at \nolinkurl{https://github.com/glandfried/tennis_atp/releases/download/atp/history.csv.zip}} including both single and double matches: if the column \texttt{double} has the letter \texttt{t}, the game is a double match.}
\es{La información contenida en los archivos \texttt{csv} contiene tanto partidas simples como dobles, indicado por la columna \texttt{double}.}
%
\en{Each game has a unique identifier (i.e. \texttt{match\_id}) and tournament's round number (i.e. \texttt{round\_number}) where 0 represents the final game, 1 the semi-final, and so on.}
\es{Tenemos un identificador por la partida (\texttt{match\_id}) y el número de ronda (\texttt{round\_number}) donde 0 representa la partida final, 1 la semi-final, etc.}
%
\en{The file also contains four columns for the players' identifiers in the game (the last two players are null in the case of a single): for example \texttt{w2\_id} is the identifier of the second player of the winning team and \texttt{l1\_name} is the name of the first player of the losing team.}
\es{Contiene cuatro columnas correspondiente a los identificadores de los jugadores y cuatro columnas correspondiente a sus nombres, por ejemplo \texttt{w2\_id} es el identificador del segundo jugador del equipo ganador (los segundos jugadores tienen valores nulos en partidas simples) y \texttt{l1\_name} es el nombre del primer jugador del equipo perdedor.}
%
\en{Finally, we have the name of the tournament (\texttt{tour\_name}), its identifier (\texttt{tour\_id}), the start date of the tournament (\texttt{time\_start}) and the type of surface (\texttt{ground}).}
\es{Finalmente tenemos el nombre del torneo (\texttt{tour\_name}), su identificado (\texttt{tour\_id}), la fecha de inicio del torneo (\texttt{time\_start}) y el tipo de piso (\texttt{ground}).}
%
\en{Here we only show the \proglang{Julia} version because it is more efficient than \proglang{Python} and \proglang{R} versions, which can be also found in the appendix \ref{sec:appendix_atp_code}.}
\es{Aquí mostramos solo la versión de \proglang{Julia} debido a que es la más eficiente, las versiones de \proglang{Python} y \proglang{R} se pueden entonrar en el anexo \ref{sec:appendix_atp_code}.}
%
\begin{lstlisting}[backgroundcolor=\color{julia},label=lst:atp, caption={\en{The history of the Association of Tennis Professionals}\es{Historia sintética}}, belowskip=0.1cm, aboveskip=0.1cm]
using CSV; using Dates
data = CSV.read("atp.csv")

>>>>

\en{The normalized posterior is found by dividing the right hand with the evidence $P(r)$.}
\es{Donde el posterior normalizado se obtiene dividiendo el lado derecho con la evidencia, $P(r)$.}

@@

\es{Es interesante notar las similitudes y diferencias entre la verosimilitud y la evidencia.}

@@

\en{In this last example we use the entire history of matches of the Association of Tennis Professionals (ATP).}
\es{En este último ejemplo utilizamos toda la historia de partidas de la Asociación de Tenistas Profesionales (ATP).}
%
\en{The database has \num{447000} games starting at year 1915 until 2020 with more than \num{19000} participating players and is publicly available\footnote{Source data at \nolinkurl{https://github.com/JeffSackmann/tennis_atp}}.}
\es{La base de datos cuenta con \num{447000} mil partidas que van desde el año 1915 hasta el 2020, en las que participan \num{19000} jugadores\footnote{Fuente de datos\nolinkurl{https://github.com/JeffSackmann/tennis_atp}}.}
%
\en{The information is formatted and stored in a single \texttt{csv} file\footnote{Available at \nolinkurl{https://github.com/glandfried/tennis_atp/releases/download/atp/history.csv.zip}} including both single and double matches: if the column \texttt{double} has the letter \texttt{t}, the game is a double match.}
\es{La información contenida en los archivos \texttt{csv} contiene tanto partidas simples como dobles, indicado por la columna \texttt{double}\footnote{Disponible en \nolinkurl{https://github.com/glandfried/tennis_atp/releases/download/atp/history.csv.zip}}.}
%
\en{Each game has a unique identifier (i.e. \texttt{match\_id}) and tournament's round number (i.e. \texttt{round\_number}) where 0 represents the final game, 1 the semi-final, and so on.}
\es{Tenemos un identificador por la partida (\texttt{match\_id}) y el número de ronda (\texttt{round\_number}) donde 0 representa la partida final, 1 la semi-final, etc.}
%
\en{The file also contains four columns for the players' identifiers in the game (the last two players are null in the case of a single): for example \texttt{w2\_id} is the identifier of the second player of the winning team and \texttt{l1\_name} is the name of the first player of the losing team.}
\es{Contiene cuatro columnas correspondiente a los identificadores de los jugadores y cuatro columnas correspondiente a sus nombres, por ejemplo \texttt{w2\_id} es el identificador del segundo jugador del equipo ganador (los segundos jugadores tienen valores nulos en partidas simples) y \texttt{l1\_name} es el nombre del primer jugador del equipo perdedor.}
%
\en{Finally, we have the name of the tournament (\texttt{tour\_name}), its identifier (\texttt{tour\_id}), the start date of the tournament (\texttt{time\_start}) and the type of surface (\texttt{ground}).}
\es{Finalmente tenemos el nombre del torneo (\texttt{tour\_name}), su identificado (\texttt{tour\_id}), la fecha de inicio del torneo (\texttt{time\_start}) y el tipo de piso (\texttt{ground}).}
%
\en{Here we only show the \proglang{Julia} version because it is more efficient than \proglang{Python} and \proglang{R} versions, which can be also found in the appendix \ref{sec:appendix_atp_code}.}
\es{Aquí mostramos solo la versión de \proglang{Julia} debido a que es la más eficiente, las versiones de \proglang{Python} y \proglang{R} se pueden entonrar en el anexo \ref{sec:appendix_atp_code}.}
%
\begin{lstlisting}[backgroundcolor=\color{julia},label=lst:atp, caption={\en{The history of the Association of Tennis Professionals}\es{Historia sintética}}, belowskip=0.1cm, aboveskip=0.1cm]
using CSV; using Dates
data = CSV.read("atp.csv")
